---
title: "knn_example1_imprvd"
author: "Vijayakumar Jawaharlal"
date: "April 23, 2014"
output: html_document
---

This is continuation of example 1. In this script, we want to implement few things to reduce "mean absolute error"

* Choosing the sample random and include all available features
* Using normalization instead of z-core standardization (used in example1) 

Here is the function for normalization

```{r}
normalize <- function(x) {
    norm <- ((x - min(x))/(max(x) - min(x)))
    return (norm)
}


marketSample <- function(smarkt, test_pct = 20) {
    train_pct = 100 - test_pct
    #Seeding for reproducibility
    set.seed(test_pct)
    test_indx <- sample(nrow(smarkt), nrow(smarkt) * test_pct/100, replace=FALSE)
    test.all <- smarkt[test_indx,]
    
    test.data <- test.all[,-9]
    test.direction <- test.all$Direction
    
    #Prep. training data  which is noting but remaing test data
    train.all <- smarkt[-test_indx,]
    train.data <- train.all[,-9]
    train.direction <- train.all$Direction
    
    rtrn_list <- list(testdata = test.data, testdir = test.direction, traindata = train.data, traindir = train.direction)
    
    return(rtrn_list)
    
}
```

Data prep using normalize function

```{r}
library(ISLR)
input.norml <- as.data.frame(lapply(Smarket[,-9],normalize))
input.norml$Direction <- Smarket$Direction

temp_list <- marketSample(input.norml, 20)
train.data <- temp_list$traindata
train.dir <- temp_list$traindir

test.data <- temp_list$testdata
test.dir <- temp_list$testdir

rm(temp_list)

```

Quickly evaluating knn algorithm

```{r}
library(class)
knn.pred <- knn(train = train.data,test = test.data,cl = train.dir,k = 3)
table(knn.pred,test.dir)
mean(knn.pred == test.dir)
```

Above 80% cases are correctly predicted which is much greater than example1. This model is doing better. Will continue working on this tomorrow.

