---
title: "kNN Using caret R package"
author: "Vijayakumar Jawaharlal"
date: "April 29, 2014"
output: html_document
---

Recently I've got familiar with caret package. Caret is a great R package which provides general interface to nearly 150 ML algorithms. It also provides great functions to sample the data (for training and testing), preprocessing, evaluating the model etc.,

To get familiar with caret package, please check following URLs

https://www.youtube.com/watch?v=7Jbb2ItbTC4
http://http://caret.r-forge.r-project.org
http://cran.r-project.org/web/packages/caret/vignettes/caret.pdf

I am going to use same dataset from previous examples. Intention of this excercise is to get familiar with caret package

##Sampling

```{r}
library(ISLR)
library(caret)

set.seed(300)
#Spliting data as training and test set. Using createDataPartition() function from caret
indxTrain <- createDataPartition(y = Smarket$Direction,p = 0.75,list = FALSE)
training <- Smarket[indxTrain,]
testing <- Smarket[-indxTrain,]

#Checking distibution in origanl data and partitioned data
prop.table(table(training$Direction)) * 100
prop.table(table(testing$Direction)) * 100
prop.table(table(Smarket$Direction)) * 100
```

creteDataParition function creates sample very effortlessly. We don't need to write complex function like previous example

##Preprocessing

kNN requires variables to be normalized or scaled. caret provides facility to preprocess data. I am going to choose centring and scaling

```{r}
trainX <- training[,names(training) != "Direction"]
preProcValues <- preProcess(x = trainX,method = c("center", "scale"))
preProcValues
```

##Training and train control

```{r}
set.seed(400)
ctrl <- trainControl(method="repeatedcv",repeats = 3) #,classProbs=TRUE,summaryFunction = twoClassSummary)
knnFit <- train(Direction ~ ., data = training, method = "knn", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 20)

#Output of kNN fit
knnFit

#Plotting yields Number of Neighbours Vs accuracy (based on repeated cross validation)
plot(knnFit)

knnPredict <- predict(knnFit,newdata = testing )
#Get the confusion matrix to see accuracy value and other parameter values
confusionMatrix(knnPredict, testing$Direction )
mean(knnPredict == testing$Direction)

#Now verifying 2 class summary function

ctrl <- trainControl(method="repeatedcv",repeats = 3,classProbs=TRUE,summaryFunction = twoClassSummary)
knnFit <- train(Direction ~ ., data = training, method = "knn", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 20)

#Output of kNN fit
knnFit

#Plotting yields Number of Neighbours Vs accuracy (based on repeated cross validation)
plot(knnFit, print.thres = 0.5, type="S")

knnPredict <- predict(knnFit,newdata = testing )
#Get the confusion matrix to see accuracy value and other parameter values
confusionMatrix(knnPredict, testing$Direction )
mean(knnPredict == testing$Direction)

```

Trying to plot ROC curve to check specificity and sensitivity

```{r}
library(pROC)
knnPredict <- predict(knnFit,newdata = testing , type="prob")
knnROC <- roc(testing$Direction,knnPredict[,"Down"], levels = rev(testing$Direction))
knnROC
plot(knnROC, type="S", print.thres= 0.5)
```


##Applying Random Forest to see the performance improvement

```{r}
set.seed(400)
ctrl <- trainControl(method="repeatedcv",repeats = 3) #,classProbs=TRUE,summaryFunction = twoClassSummary)

# Random forrest
rfFit <- train(Direction ~ ., data = training, method = "rf", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 20)

rfFit

plot(rfFit)

rfPredict <- predict(rfFit,newdata = testing )
confusionMatrix(rfPredict, testing$Direction )
mean(rfPredict == testing$Direction)

#With twoclasssummary
ctrl <- trainControl(method="repeatedcv",repeats = 3,classProbs=TRUE,summaryFunction = twoClassSummary)
# Random forrest
rfFit <- train(Direction ~ ., data = training, method = "rf", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 20)

rfFit

plot(rfFit)

#Trying plot with some more parameters
plot(rfFit, print.thres = 0.5, type="S")

rfPredict <- predict(rfFit,newdata = testing )
confusionMatrix(rfPredict, testing$Direction )
mean(rfPredict == testing$Direction)


```

Ploting ROC curve

```{r}
library(pROC)
rfPredict <- predict(rfFit,newdata = testing , type="prob")
rfROC <- roc(testing$Direction,rfPredict[,"Down"], levels = rev(testing$Direction))
rfROC
plot(rfROC, type="S", print.thres= 0.5)
```
#Yet to learn about interpreting ROC curve :)