---
title: "Wisconsin Breast Cancer Diagnostic Data Set"
author: "Vijayakumar Jawaharlal"
date: "May 3, 2014"
output: html_document
---

This is the sample data set used in "Machine Learning with R" book to teach kNN algorithm.

##Data exploration

```{r}
library(caret)
setwd("/Users/njvijay/big_data/Github/MachineLearning-R/knn/data")
srcData <- read.csv("wisc_bc_data.csv")
str(srcData)
#Removing id field which is not required for prediction.
wcData <- srcData[,-1]
table(wcData$diagnosis)
prop.table(table(wcData$diagnosis))

```

It has about 569 observations(biopsies) with 32 variables. diagnosis is the response variable with 2 levels B-begnin and M - melignant. All of the predictors are quantitative with different measurement. These predictors needs to be standardize before applying kNN algorithm.

##Data Spliting
Reserve 20% data for validation. 80% for cross validation and model building
```{r}
set.seed(456)
idx <- createDataPartition(wcData$diagnosis,p = 0.8,list = FALSE)
train <- wcData[idx,]
test <- wcData[-idx,]
#Checking proportion of data spliting
prop.table(table(train$diagnosis))
prop.table(table(test$diagnosis))
```
% is maintained in the data spliting

##Setting parallelism

```{r}
library(doMC)
registerDoMC(cores = 3)

```

Building the model

```{r}
set.seed(400)
ctrl <- trainControl(method="repeatedcv",repeats = 3,classProbs=TRUE,summaryFunction = twoClassSummary)
knnFit <- train(diagnosis ~ ., data = train, method = "knn", trControl = ctrl, preProcess = c("center","scale"), tuneLength = 31)
knnFit
plot(knnFit)
```

Validate the model on the test data

```{r}
knnPredict <- predict(knnFit,newdata = test)
#Get the confusion matrix to see accuracy value and other parameter values
confusionMatrix(knnPredict, test$diagnosis )
```

Measuring performance using ROCR package

```{r}
knnperfPredict <- predict(knnFit,newdata = test,type="prob")
knnperfPredictB <- knnperfPredict$B
#Creating dummy vars
testdiagnosis <- ifelse(test$diagnosis == 'B', 1,0)

library(ROCR)
perfprod <- prediction(knnperfPredictB, testdiagnosis)
perf <- performance(perfprod,"tpr","fpr")
plot(perf)

perf <- performance(perfprod,"sens","spec")
plot(perf)

```

ROCR supports only binary classification. It is a limitation.






